---
title: "UFO Sightings Exploratory Analysis "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, include=FALSE}

library(mapview)
library(sf)
library(tidyverse)
library(here)
library(ggplot2)
library(ggmap)
library(rworldmap)
library(readxl)
library(plotly)
library(pander)
library(usmap)
library(gridExtra)

here::here()

data <- read_csv(here("nuforc_reports.csv"))

```

### Maps of Sightings


```{r map, echo=FALSE, warning=FALSE, message=FALSE}

longitudes <- na.omit(data$city_longitude) # Get rid of N/A rows 
latitudes <- na.omit(data$city_latitude) # Get rid of N/A rows

newrows <- data.frame(longitudes,latitudes) # Create new data frame with our revised variables

world <- getMap(resolution = "medium") # Get data for map

# Make a map for all UFO sightings

ggplot() +
    geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = NA, colour = "black") + # Plot Map
    geom_point(data = newrows, aes(x = longitudes, y = latitudes) , colour = "green4", cex = .05 ) + # Plot Data
    theme_classic() + # Remove Grey Background
    ylim(10, 70) +
    xlim(-170,-50) +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("UFO Sightings")+
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))


```


##### As we can see from the map, we have a lot of sightings. We are going to look at the relationship between the year and the number of sightings to help us determine a place to split the data. We believe a dichotomous data set will be more useful in our analysis.  


```{r, echo=FALSE, warning=FALSE, message=FALSE}

```{r, echo=FALSE, warning=FALSE, message=FALSE}

data$sighting_year <- lubridate::year(data$date_time)
year <- data %>% group_by(sighting_year) %>% summarise(count = n()) %>% arrange(desc(count))


ggplot(year, aes(x = sighting_year, y = count)) + geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .35)) +
  xlab("Sighting Year") +
  ylab("Count")

```

##### We can see from this chart that there is a huge jump in sightings after 2005. We are unsure as to why this jump occurs, but we believe this is a good place to divide the data. We are going to call the sightings before 2005 "Early Data" and we are going to call the sightings after 2005 "Later Data". 


### Maps of Split Data

```{r, echo=FALSE, warning=FALSE, message=FALSE}


data$sighting_year <- lubridate::year(data$date_time) 


# Create Clean Data Table 


newdata <- data.frame(data$sighting_year ,data$city_longitude, data$city_latitude)
newdata2 <- na.omit(newdata) # Get rid of empty rows
newdata2$sighting_year <- newdata2$data.sighting_year
newdata2$longitude <- newdata2$data.city_longitude
newdata2$latitude <- newdata2$data.city_latitude

newdata2$data.sighting_year <- NULL
newdata2$data.city_longitude <- NULL
newdata2$data.city_latitude <- NULL


# Split


earlydata <- subset(newdata2, newdata2$sighting_year <= 2005) # UFO sightings before 2005
laterdata <- subset(newdata2, newdata2$sighting_year > 2005) # UFO sightings after 2005 


# Make a map for earlydata


(mapufo <- ggplot() +
    geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = NA, colour = "black") + # Plot Map
    geom_point(data = earlydata, aes(x = earlydata$longitude , y = earlydata$latitude) , colour = "red", cex = .05 ) + # Plot Data
    theme_classic() + # Remove Grey Background
    ylim(10, 70) +
    xlim(-170,-50) +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Early Data"))+
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# Make a map for laterdata


(mapufo <- ggplot() +
    geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = NA, colour = "black") + # Plot Map
    geom_point(data = laterdata, aes(x = laterdata$longitude , y = laterdata$latitude) , colour = "blue", cex = .05 ) + # Plot Data
    theme_classic() + # Remove Grey Background
    ylim(10, 70) +
    xlim(-170,-50) +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Later Data"))+
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

##### We can see from these maps that a lot of the sightings tend to group together in certain areas. What is the reason for this? Our first perdiction is that there is a correlation between UFO activity and the U.S. nuclear power plants. We will be exploring this hypothesis in the next section.

### Nuclear Power Plant Analysis



```{r, echo=FALSE, warning=FALSE, message=FALSE}


NuclearData <- read_excel(here("NuclearData.xlsx"))

NuclearData2 <- data.frame(NuclearData$`Reactor Name`,NuclearData$Longitude, NuclearData$Latitude)
NuclearData2$name <- NuclearData2$NuclearData..Reactor.Name.
NuclearData2$longitude <- NuclearData2$NuclearData.Longitude
NuclearData2$latitude <- NuclearData2$NuclearData.Latitude
NuclearData2$NuclearData..Reactor.Name. <- NULL
NuclearData2$NuclearData.Longitude <- NULL
NuclearData2$NuclearData.Latitude <- NULL

NuclearData3 <- na.omit(NuclearData2)


# Going to plot the nuclear power plants and the ufo sightings on the same map


MainStates <- map_data("state")

mapufo <- ggplot() +
    geom_polygon(data = MainStates, aes(x = long, y = lat, group = group), fill = NA, color = "black") + # Plot Map
    geom_point(data = earlydata, aes(x = earlydata$longitude , y = earlydata$latitude, text = paste0(earlydata$sighting_year)) , color = "red", cex = .3 ) + # Plot Data
    geom_point(data = NuclearData3, aes(x = NuclearData3$longitude , y = NuclearData3$latitude, text = paste0(NuclearData3$name)) , color = "blue", cex = 1) +
    theme_classic() + # Remove Grey Background
    ylim(25, 50) +
    xlim(-125,-65) +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("UFO Sightings Before 2005 With U.S. Nuclear Power Plants") +
    theme(plot.title = element_text(hjust = 0.5 , face = "bold"))

fig2 <- ggplotly(mapufo, tooltip = "text")

fig2

```

##### From the map alone we can start to see that there is more sightings near power plants. 

##### In this next section I am going to compare the number of power plants per state vs the number of sightings per state and average population. I am going to be working with the early data (before 2005) and I will be using the 2000 census for our population data. 

```{r states, echo=FALSE, warning=FALSE, message=FALSE}


# Modifying Our Data


NucPerState <- read_excel(here("NucPerState.xlsx")) # Data set w/states and their count of power plants

statedata <- data.frame(data$sighting_year , data$state) # Creating a data set with just the year/state
statedata2 <- na.omit(statedata) # Get rid of empty rows

statedata2$sighting_year <- statedata2$data.sighting_year
statedata2$state <- statedata2$data.state
statedata2$data.sighting_year <- NULL
statedata2$data.state <- NULL


earlydata2 <- subset(statedata2, statedata2$sighting_year <= 2005) # UFO sightings before 2005
earlydata22 <- earlydata2 %>% group_by(state) %>% summarise(count = n()) # This counts the number of sightings per state


earlydata22$State <- earlydata22$state
perstate1 <- left_join(earlydata22, NucPerState, by = "State") # Note left_join is better than cbind
perstate1$Population <- as.integer(perstate1$Population) # Converting Population from string values to int



# Creating and Displaying Table


summary <- perstate1 %>% group_by(Count) %>% summarise(avg_sightings = mean(count), avg_pop = mean(Population))
pander(summary)


# Making Plots


plot1 <- ggplot() +
  geom_point(data = summary, aes(x = summary$Count, y = summary$avg_sightings), color = "red", cex = 2) +
    xlab("Number of Power Plants in the State") +
    ylab("Average # of Sightings")

plot2 <- ggplot() +
  geom_point(data = summary, aes(x = summary$Count, y = summary$avg_pop), color = "blue", cex = 2) +
    xlab("Number of Power Plants in the State") +
    ylab("Average Population")

grid.arrange(plot1, plot2)

```

##### Based on our visuals it seems like we have a positive correlation between the number of power plants, the number of sightings, and the population. 

##### It's important to note that California is our outlier of this data. California has significantly more sightings and is the most populated state. Based on our "README" I think California has some other explanatory variables (San Andreas Fault, Los Angeles, Earthquakes) that contribute to more sightings. We will have to consider this as we continue our analysis. 

##### We consider the population to be our cofounding variable as it influences the number of sightings as well as the number of power plants in a state. We will be using some regression analysis to determine the relationship between population, sightings, and nuclear power plants. 

#### Scaled Down Population

<<<<<<< HEAD
<<<<<<< HEAD
##### I tried to plot everything on one graph, but it didn't turn out great because sightings are in the hundreds and population is in the millions. I have decided to scale down population by dividing all our values by the population of the state with the lowest population value. In this case the state with the lowest population is Wyoming with 493,782.
=======

##### I tried to plot everything on one graph, but it didn't turn out great because sightings are in the hundreds and population is in the millions. I have decided to scale down population by dividing all our values by the population of the state with the lowest population value. In this case the state with th elowest population is Wyoming with 493,782.
>>>>>>> 76e359c3f75bab3e7d587c0d30f70ad262ec1c71
=======

##### I tried to plot everything on one graph, but it didn't turn out great because sightings are in the hundreds and population is in the millions. I have decided to scale down population by dividing all our values by the population of the state with the lowest population value. In this case the state with th elowest population is Wyoming with 493,782.

>>>>>>> 67457278a9e25921b3ba479c82de2171a9bd7cd8


```{r, echo=FALSE, warning=FALSE, message=FALSE}

earlydata22$State <- earlydata22$state
perstate1 <- left_join(earlydata22, NucPerState, by = "State")
perstate1$Population <- as.integer(perstate1$Population)

perstate1$Population <- perstate1$Population / 493782


scaled_summary <- perstate1 %>% group_by(Count) %>% summarise(avg_sightings = mean(count), avg_pop_scaled = mean(Population))

#pander(scaled_summary)

(ggplot() +
  geom_point(data = scaled_summary, aes(x = scaled_summary$Count, y = scaled_summary$avg_pop_scaled), color = "blue", cex = 2) +
    geom_point(data = scaled_summary, aes(x = scaled_summary$Count, y = scaled_summary$avg_sightings), color = "red", cex = 2) +
    xlab("Number of Power Plants in the State") +
    ylab("Average Scaled Population and Average Sightings"))

```

##### I will now try and perform linear regression with number of power plants and populatoin as our independent variables, and sightings as our dependent variable. 

<<<<<<< HEAD
<<<<<<< HEAD
=======

>>>>>>> 67457278a9e25921b3ba479c82de2171a9bd7cd8
```{r, echo=FALSE, warning=FALSE, message=FALSE}

model1 <- lm(scaled_summary$avg_sightings ~ scaled_summary$avg_pop_scaled + scaled_summary$Count )
summary(model1)

#cor(scaled_summary$avg_sightings, scaled_summary$avg_pop_scaled, method = "pearson")
par(mfrow = c(2,2))
plot(model1)
```

##### I believe that we are unable to perform a linear regression analysis on this data because the data isn't linear. 


<<<<<<< HEAD
=======
>>>>>>> 76e359c3f75bab3e7d587c0d30f70ad262ec1c71
=======
>>>>>>> 67457278a9e25921b3ba479c82de2171a9bd7cd8
#### Nuclear Range 


##### One degree of latitude is approxiamtely 69 miles, and one degree of longitude is about 54.6 miles. We are going to be looking at the number of sightings that occur within one degree from our power plants. This accounts for about 15,069.6 square miles.   

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Creating a new table with ranges for our longitudes and latittudes


NuclearData4 <- data.frame(NuclearData3$name, NuclearData3$longitude - 1, NuclearData3$longitude + 1, NuclearData3$latitude - 1, NuclearData3$latitude + 1)
colnames(NuclearData4) = c("name", "l_longitude", "u_longitude", "l_latitude", "u_latitude")


# Iterating through both data sets to count the number of sightings in our ranges 


x = c()

for(i in 1:nrow(NuclearData4)){
    y <- 0
   for (j in 1:nrow(earlydata)){
       if(NuclearData4[i,2] <= earlydata[j,2] && earlydata[j,2] <= NuclearData4[i,3] && NuclearData4[i,4] <= earlydata[j,3] && earlydata[j,3] <= NuclearData4[i,5]) {y <- y + 1}
       else {y <- y+0}
       } 
    if(y >= 0){ x <- append(x, y)}
} 
  

```




```{r echo=FALSE, warning=FALSE, message=FALSE}

counting <- data.frame(x) # Making vector into data frame so its easier to put into table


# Table with Reactor name and the number of sightings near


newtable <- cbind(NuclearData4$name, counting) 
colnames(newtable) = c("Reactor Name", "Count of Sightings Near")
#pander(newtable)


# Table that counts the total number of sightings near Vs not near


r = rbind( sum(counting), 4737 - sum(counting),  4737)
q = rbind("84%", "16%", "")
p = cbind(r,q)
colnames(p) = c("Number of Sightings", "Ratio")
rownames(p)= c("Near Power Plants", "Not Near" ,"Total")

pander(p)

```

##### From this data we can see that about 84% of our sightings occur within about 55-70 miles from our power plants. 

##### I have decided to zoom in on the map and check the count for two of our power plants. I am using the power plant "Diablo Canyon 1" which on the table has 14 counts, but as you will see on the map there are only 10 sightings in our range. "Rancho Seco" has 85 counts on the table, but only 45 on the map. I am unsure as to why the map isn't corresponding to the table. 


```{r echo=FALSE, warning=FALSE, message=FALSE}


fig <- (mapufo <- ggplot() +
    geom_point(data = earlydata, aes(x = earlydata$longitude , y = earlydata$latitude, text = paste0(earlydata$sighting_year)) , color = "red", cex = 1 ) + # Plot Data
    geom_point(data = NuclearData3, aes(x = NuclearData3$longitude , y = NuclearData3$latitude, text = paste0(NuclearData3$name)) , color = "blue", cex = 2 ) +
   geom_segment(aes(x = -121.85554, y = 	34.21157, xend = 	-121.85554, yend = 	36.21157)) +
   geom_segment(aes(x = -121.85554, y = 	36.21157, xend = 	-119.85554, yend = 	36.21157)) +
   geom_segment(aes(x = -119.85554, y = 	36.21157, xend = 	-119.85554, yend = 	34.21157)) +
   geom_segment(aes(x = -119.85554, y = 	34.21157, xend = 	-121.85554, yend = 	34.21157)) +
     
   geom_segment(aes(x = -122.12167, y = 	37.34528, xend = 	-122.12167, yend = 	39.34528)) +
   geom_segment(aes(x = -122.12167, y = 	39.34528, xend = 	-120.12167, yend = 	39.34528)) +
   geom_segment(aes(x = -120.12167, y = 	39.34528, xend = 	-120.12167, yend =  37.34528)) +
   geom_segment(aes(x = -120.12167, y = 	37.34528, xend = 	-122.12167, yend = 	37.34528)) +
    theme_classic() + # Remove Grey Background
    ylim(30, 40) +
    xlim(-123, -118) +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("UFO Sightings at Diablo Canyon and Rancho Seco Power Plants")) +
    theme(plot.title = element_text(hjust = 0.5 , face = "bold"))

fig3 <- ggplotly(fig, tooltip = "text")

fig3
```